### 2021-08-21 doc/thesis/chapters/implementation.tex
impl vhdl done

The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter~\ref{cha:SystemArchitectureAndConcept}.\\

Figure~\ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
The actual simulator interfaces, as described in Section~\ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\

\includepicture [1.0] [0]  {Block diagram of the implemented testbench architecture.} {bd_testbench_architecture_vhdl} {img/draw.io/bd_testbench_architecture_vhdl}

The test cases are using the external verification data, that was previously generated by the Matlab model.
This data is supplied to the input of the DUT.
The DUTs' input and output is implemented as an AXI Stream interface, which is an interface that features a valid-, as well as a ready-signal along with the data.
The cocotb framework supports both an AXI Stream Master, and an AXI Stream Slave class, to write and read the input and output, respectively.
In that way, the interface signals are abstracted to simple read- and write-function calls in the Python code, while the actual signal handling is done in the mentioned cocotb library classes.
A clock generator class is also available from cocotb, which is used to provide the system clock to the DUT.\\

The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots with Pythons' matplotlib library.
The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
The class called \textit{fm\_receiver\_model} serves as the golden model reference that is used to compare the DUT against.
Internally, it loads the Matlab verification data.
This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as the single source of truth in this project.\\

At the end of the test cases, the DUTs' output data is written to a file.
This data is later used to directly compare the different implementation variants against each other.

@@

Common strategies for testing are unit tests and integration tests, which represent different levels of testing.
Unit tests specifically target a single unit, while integration tests verify the combination of multiple units as a whole.
Both of these stragegies are important.
Given all unit tests being successful, it is ensured that all these single units are implemented correctly.
However, their combination and interaction may still be subject to error -- this is where integration tests unfold their potential.
Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.\\

The testbench architecture presented above represents an integration test, since the DUT is tested as an entire block.
The advantage thereof is, that it tests the DUT in the same way as it will be integrated within the target hardware, the FPGA.
However, in case the DUT fails to produce a correct output in the integration test, it is not always visible where exactly the error is located.
The source of error may be hidden anywhere within the nested and interconnected units.
Thus, in addition to the DUTs' interfaces, further signals are logged from the inside of the DUT.
These internal signals can be accessed through the Python testbench.\\

In this context, there are further keywords that need to be mentioned -- black box testing and white box testing.
These are testing strategies, and describe how the DUT is treated.
In the case of black box, the DUTs internals, such as implementation details, are not known.
Thus, only the exposed interfaces can be tested.
In the white box as a contrary, every detail of the internal structure is known.
Thus, even single signals can be accessed, which is what is done in this Python testbench.
The advantage of the white box is, that these internals can be tested.
However, any time the implementation changes, even if it is just the naming of a signal, the testbench code needs to be adapted as well.
In black box testing, the testbench code is independent of the detailed implementation of the DUT, as long as it maintains its required behaviour at the exposed interfaces.
However, the disadvantage is that internal states cannot be observed.\\

The Python testbench that is implemented combines the advantages of both worlds and uses black box testing, as well as white box testing.

@@

In this project, the only module that has its dedicated unit test is the FIR filter.
The reason therefor is, that it is the most complex unit within the entire design and is used in multiple instances.
All the other design units mostly consist of simple delay lines, additions or substractions and thus a special unit test is not implemented.

@@

In the VHDL version of the FM radio receiver IP, this handshaking is not implemented according to the AXI stream specification.
Specifically, the ready signal handshake is implemented with a workaround, to simplify the development.
The DSP chain of the IP is not fully pipelined and thus cannot consume an input sample on every clock cycle.
This is mainly because of the FIR filter implementation -- it takes multiple clock cycles to compute an output sample, before it can take a new sample.
Ideally, this situation should be handled by the ready signal, since this is exactly what it is intended for.
However, in order to simplify the implementation a strobe generator is used, which is a counter that generates a repetitive pulse with a defined period.
This pulse, in combination with some additional logic, is used as the ready signal, to guarantee enough spacing between two consecutive input samples.

@@

The VHDL code for this register interface is generated by a Python script which is named \textit{Register Engine}.
The registers can be defined in a YAML-file.
Additionally, documentation in the form of a Markdown language file is generated, and -- more importantly -- a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\

>>>>

The architecture for this testbench is created according to the guidelines and assumptions that are developed in Chapter~\ref{cha:SystemArchitectureAndConcept}.\\

Figure~\ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
The split between the Python environment and the VHDL implementation is clearly visible and is represented from the developer's perspective.
The actual simulator interfaces, as described in Section~\ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain a better overview.\\

\includepicture [1.0] [0]  {Block diagram of the implemented testbench architecture.} {bd_testbench_architecture_vhdl} {img/draw.io/bd_testbench_architecture_vhdl}
%TODO: remove the yellow color

The test cases are using the verification data, that was previously generated by the Matlab model.
This data is supplied to the input of the DUT.
The DUT's input and output is implemented as an AXI Stream interface, which is an interface that features a valid-, as well as a ready-signal along with the data.
The cocotb framework supports both an AXI Stream Master, and an AXI Stream Slave class, to write and read the input and output, respectively.
In that way, the interface signals are abstracted to simple read- and write-function calls in the Python code, while the actual signal handling is done in the mentioned cocotb library classes.
A clock generator class is also available from cocotb, which is used to provide the system clock to the DUT.\\

The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as the comparison against defined error thresholds or the creation of graphical plots with Python's library matplotlib.
The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
The class called \textit{fm\_receiver\_model} serves as the golden model reference that is used to compare the DUT against.
Internally, it loads the Matlab verification data.
This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is seen as the single source of truth in this project.\\

At the end of the test cases, the DUT's output data is written to a file.
This data is later used to directly compare the different implementation variants against each other.

@@

Common strategies for testing are unit tests and integration tests, which represent different levels of testing.
Unit tests specifically target a single unit, while integration tests verify the combination of multiple units as a whole.
Both of these strategies are important.
Given all unit tests being successful, it is ensured that all these single units are implemented correctly.
However, their combination and interaction may still be subject to error -- this is where integration tests unfold their potential.
Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be revealed with these integration tests.\\

The testbench architecture presented in Figure~\ref{fig:bd_testbench_architecture_vhdl} represents an integration test, since the entire DUT is tested as a single block, with all the single units already being integrated.
The advantage thereof is, that it tests the DUT in the same way as it will be integrated within the target hardware, the FPGA.
However, in case the DUT fails to produce a correct output in the integration test, it is not always visible where exactly the error is located.
The source of error may be hidden anywhere within the nested and interconnected units.
Thus, in addition to the DUT's interface signals, further signals from the inside of the DUT are logged.
These internal signals can be accessed through the Python testbench.\\

In this context, there are further keywords that need to be mentioned -- black box testing and white box testing.
They are testing strategies, and describe how a DUT is treated in a test procedure.
In the case of black box testing, the DUT's internals, such as implementation details, are not known.
Thus, only the exposed interfaces can be tested.
In white box testing, in contrast, every detail of the internal structure is known.
Thus, any single internal signal can be accessed, which is what is done in this Python testbench.
The advantage of the white box is, that these internals can be tested.
However, any time the implementation changes, even if it is just the naming of a signal, the testbench code needs to be adapted as well.
In black box testing, the testbench code is independent of the detailed implementation of the DUT, as long as it maintains its defined behaviour at the exposed interfaces.
However, the disadvantage is that internal states can not be observed directly.\\

The Python testbench that is implemented, combines the advantages of both worlds and uses black box testing, as well as white box testing.

@@

In this project, the only module that has its dedicated unit test is the FIR filter.
The reason therefor is, that it is the most complex unit within the entire design and is used in multiple instances.
All the other design units mostly consist of simple delay lines, additions or substractions and thus, special unit tests are not implemented there.

@@

In the VHDL version of the FM radio receiver IP, this handshaking is not implemented according to the AXI stream specification.
Specifically, the ready signal handshake is implemented with a workaround, to simplify the development.
The DSP chain of the IP is not fully pipelined and thus cannot consume an input sample on every clock cycle.
This is mainly because of the FIR filter implementation -- it takes multiple clock cycles to compute an output sample, before it can take a new sample.
Ideally, this situation should be handled by the ready signal, since this is exactly what it is intended for.
However, in order to simplify the implementation, a strobe generator is used, which is a counter that generates a repetitive pulse with a defined period.
This pulse, in combination with some additional logic, is used as the ready signal, to guarantee enough spacing between two consecutive input samples and therefore limit the input dataflow.

@@

The VHDL code for this register interface is generated by a Python script which is named \textit{Register Engine}.
The registers can be defined in a YAML file.
Additionally, documentation in the form of a Markdown language file is generated, and -- more importantly -- a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\
